 ---
title: "TFBS_visualization"
author: "Ciera Martinez"
date: "July 10, 2017"
output: pdf_document
---

**Purpose**: The purpose is to prototype 
1. what type of analysis I can do with the data set I have created. 
2. see what else in terms of information / data I need to collect.

### Set up 
```{r}
## Libraries
library(ggplot2)
library(dplyr)
library(stringr)
```

## TFBS_df

Read in and clean up all TFBS found 
```{r}
# This takes about 5 min
TFBS_df_0 <- read.table("../data/outputs/map_motif_output/all.csv", sep = "\t", header = FALSE)

# So not to have to read in again
TFBS_df <- TFBS_df_0
TFBS_df <- TFBS_df[,-1]

colnames(TFBS_df) <- c("position", "score",	"seq_len",	"species", "raw_position",	"strand",	"align_position",	"alignment_file", "motif_file")

## as.factor species
head(TFBS_df)
TFBS_df$species <- as.factor(TFBS_df$species)

## Retrieve species
## Takes about 3 min
speciesSplit  <- data.frame(do.call('rbind', strsplit(as.character(TFBS_df$species), split = "|", fixed = TRUE)))

## Remove old species id and replace
TFBS_df <- TFBS_df[,-4]
TFBS_df <- cbind(TFBS_df, speciesSplit$X6)
colnames(TFBS_df)[9] <- "species"

## Clean up alignment file notation
TFBS_df$alignment_file <- gsub("output_", "", TFBS_df$alignment_file) 
TFBS_df$alignment_file <- gsub(".fa", "", TFBS_df$alignment_file)

## Clean up motif_file
TFBS_df$motif_file <- gsub(".fm", "", TFBS_df$motif_file)

## Check
head(TFBS_df)
str(TFBS_df)
summary(TFBS_df)
```

## PhastCons score per alignment

Read in and clean up PhastCons scores.

```{r}
# takes about 2 min
cons_scores_0 <- read.csv("../data/outputs/consScore/allConsScore.csv")

## so not to reload dataset
cons_scores <- cons_scores_0

cons_scores <- cons_scores[,-1]
colnames(cons_scores)[1] <- "raw_MEMB002A_pos"
head(cons_scores)

colnames(cons_scores)[3] <- "alignment_file"
cons_scores$alignment_file <- gsub(".fa", "", cons_scores$alignment_file)
cons_scores$alignment_file <- gsub("output_", "", cons_scores$alignment_file) 
```

## Kvon data about expression in D. mel

```{r}
## Kvon data Info
kvon <- read.csv("../../montium_analyses_1_alignment/outputData/dataframe/kvonMergedWithThomasStageSpecific_15November2016.csv", header=T, na.strings=c("","NA"), check.names = TRUE)
kvon <- kvon[,-1]

colnames(kvon)[3] <- "alignment_file"
colnames(kvon)[5:9] <- c("DNase_Peaks_stage09", "DNase_Peaks_stage10", "DNase_Peaks_stage11", "DNase_Peaks_stage14", "DNase_Peaks_sumStages")
colnames(kvon)[12] <- "positive_GAL4_anyStage"
```

## Merge Data

```{r}
## Merge Kvon expression information with TFBS.
kvon_sub <- kvon[,c(3,12,13,9)]
head(kvon_sub)
length(levels(as.factor(kvon_sub$alignment_file))) #7792

TFBS_kvon_df <- merge(TFBS_df, kvon_sub, by = "alignment_file", all.x = TRUE)

head(TFBS_kvon_df)


## Merge TFBS alignment_pos and raw_pos to cons_score

head(TFBS_df)
MEMB002A_map <- subset(TFBS_df, species == "MEMB002A")


head(MEMB002A_map)

head(cons_scores)
```

## Create Subset file

```{r}

all_regions <- levels(as.factor(TFBS_kvon_df$alignment_file))
length(all_regions) # 3483

## randomly choose 200
sub_regions <- sample(all_regions, 500)
length(sub_regions)
levels(as.factor(sub_regions))

sub_TFBS_kvon_df <- TFBS_kvon_df[TFBS_kvon_df$alignment_file %in% sub_regions,]

## check
levels(as.factor(sub_TFBS_kvon_df$alignment_file))

## write.c
## write.csv(sub_TFBS_kvon_df, "../data/outputs/combined_TFBS_kvon_10August2017.csv")

##### ConsScore subfile
## TFBS_df <- read.csv("../data/outputs/combined_TFBS_kvon_10August2017.csv")
sub_regions <- levels(as.factor(TFBS_df$alignment_file))
sub_regions <- as.data.frame(sub_regions)

## write.table(sub_regions, "../data/outputs/sub_regions.csv", quote = FALSE, row.names = FALSE)

dim(cons_scores)
sub_cons_score <- cons_scores[cons_scores$alignment_file %in% sub_regions,]
dim(sub_cons_score)

## write.csv(sub_cons_score, "../data/outputs/sub_cons_score_10August2017.csv")


##########Alignment Keys

## alignment key (already subsetted previously, see quiver)
## About ~2 min
alignmentKeys <- read.table("../data/outputs/alignment_keys/sub_regions_alignment_key.csv", sep = "\t", header = TRUE)

alignmentKeys <- alignmentKeys[,-1]


## Retrieve species
## Takes about 3 min
speciesSplit  <- data.frame(do.call('rbind', strsplit(as.character(alignmentKeys$species), split = "|", fixed = TRUE)))

## Remove old species id and replace
alignmentKeys <- alignmentKeys[,-3]
alignmentKeys <- cbind(alignmentKeys, speciesSplit$X6)
colnames(alignmentKeys)[4] <- "species"

##Subset by MEMBOO2A
alignmentKeys <- subset(alignmentKeys, species == "MEMB002A")

## Clean up alignment file notation
alignmentKeys$alignment_file <- gsub("output_", "", alignmentKeys$alignment_file) 
alignmentKeys$alignment_file <- gsub(".fa", "", alignmentKeys$alignment_file)

head(alignmentKeys)

write.csv(alignmentKeys, "../data/outputs/sub_alignment_keys_11August2017.csv", row.names = FALSE)
```

Most conserved Regions.

Subsetted using shell. See quiver notebook.
```{r}
most_conserved <- read.csv("../data/outputs/sub_mostConserved.csv")



